---
title: "GCN Tutorial: Graph Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
source("./gcns.R")
```

## Data formatting
First, let's import some data. These are dummy genetic networks: six in total, each with 100 genes.
```{r include=TRUE}

networks = lapply(1:6, function(i){
  read.table(paste("./sample_gcns/region_",i,".csv",sep=""), sep=",", fill=TRUE, stringsAsFactors=FALSE, header=TRUE)
})

```

We'll start by convert our covariance matrices into graphs.
```{r}
graphs = lapply(networks, function(x){
  graph_from_adjacency_matrix(as.matrix(x), weighted=TRUE)
})
```

## Hub gene analysis
Next, we can get the top hub genes for each graph.
```{r}
top_hub_genes = lapply(graphs, getTopXPercentHubGenes, percentage=0.25)

print(top_hub_genes)

```



```{r}
#TODO: a method to count which ones are in common
```

```{r}
#TODO: a method to "impute" missing columns/rows so that the nodes are the same
```

## Modularization
```{r}
#TODO: a method for modularization based on  that algorithm I found a while ago
```


## Binary Graphs
There are a few analyses we can do with binarized graphs, i.e. two genes are either covariable or not covariable. We next binarize our networks using a cutoff of 0.75. This is a fairly lax cut-off.
```{r}
binary_mx = lapply(networks, convertAdjacencyToBinary, cutoff=0.75)

```


We can see how frequently each possible edge occurs
```{r}
counts = calculateEdgeOccurrences(binary_mx) 
table(counts$Count)
```


We can take a closer look at each unique edge network.
```{r}
binarygraphs = lapply(binary_mx, graph_from_adjacency_matrix)

unique_edge_networks = identifyUniqueEdges(binarygraphs)

```


```{r}
# TODO: statistics about unique edge networks and comparisons to original networks
```

